\chapter{Installing}

Mongrel2 is designed to build on most modern Unix systems, specifically Linux
and Mac OSX.  It is written in C (\emph{not Ruby}) and uses fairly vanilla
C and standard libraries, except for one piece that implements the internal
coroutines.  Other than this, you should be able to compile and install Mongrel2
with nothing more than \shell{make all install} after you've installed
all the dependencies.

Now, if when I said dependencies you started to groan at having to install
software to use my software, well my friend, welcome to the future.  You
said you don't want people reinventing the wheel, right?  Great, that means
you need to install software for my software to work.  It's either that or
wait 10 years for me to build everything from scratch like some arrogant
jackass.

We good now?  Great, let's get started.

\section{Install Dependencies}

To get everything working you will need the following dependencies:

\begin{itemize}
\item GNU make (gmake).
\item \href{http://zeromq.org}{ZeroMQ} for the messaging.
\item \href{http://python.org}{Python 2.6} for the m2sh management and examples.
\item \href{http://www.sqlite.org/}{SQLite3} and make sure \emph{Python} has the ability to access SQLite3.
\item \href{http://pypi.python.org/pypi/distribute}{distribute} and \href{http://pypi.python.org/pypi/pip/0.7.2}{pip}.
\item \href{http://github.com/zeromq/pyzmq}{PyZMQ} for the examples.
\item \href{http://webpy.org/}{web.py} for the examples.
\end{itemize}

If you install these things in this order, then everything should be good.
Typically people run into problems if they do this out of order; for example,
trying to install Mongrel2 before they have distribute and pip installed.

Since every system is different, it is difficult to tell you exactly how to
install required packages for your OS, but here's how I did it on my computer:


\begin{code}{Installing Dependencies on ArchLinux}
\begin{Verbatim}
# install ZeroMQ
> wget http://www.zeromq.org/local--files/area:download/zeromq-2.0.7.tar.gz
> tar -xzvf zeromq-2.0.7.tar.gz
> cd zeromq-2.0.7/
> ./configure
> make
> sudo make install

# install python
> sudo pacman -S python

# install sqlite3
> sudo pacman -S sqlite3

# install distribute and pip
> sudo pacman -S python-pip
> sudo pip install distribute

# install PyZMQ from github
> sudo pacman -S git
> git clone http://github.com/zeromq/pyzmq.git
> sudo python setup.py install

# web.py
> sudo pip install web.py
\end{Verbatim}
\end{code}

If you run into parts that your OS is missing, which is likely on
Debian and SuSE systems, then you'll have to go and figure out
how to install it.


\section{Building Mongrel2}

If everything went well you should be able to grab the Mongrel2 source
and try building it.  There's two ways you can get the source code to
Mongrel2:

\begin{enumerate}
\item Install \href{http://fossil-scm.org}{Fossil SCM} and check out the source.
\item Grab the source .zip release and install it from there.
\end{enumerate}

\subsection{Using Fossil}

Since it's early in Mongrel2 development, we recommend you get fossil and install it
so you can track the latest fixes and development.  Here's the easiest way to
do that:

\begin{enumerate}
\item Go to \href{http://www.fossil-scm.org/download.html}{the fossil download page} and
    grab the binary that fits your system, or the source tar.gz.
\item Follow their \href{http://fossil-scm.org/index.html/doc/tip/www/build.wiki}{install instructions}
    to get it on your system.
\item If you find that your version of fossil ends up having problems with mongrel2.org let us know since
    we might need to upgrade.
\end{enumerate}

Once you have fossil you can then get the Mongrel2 source and open it up:

\begin{code}{Cloning the Mongrel2 Source}
\begin{Verbatim}
> mkdir ~/fossils
> fossil clone http://mongrel2.org ~/fossils/mongrel2.fossil
> mkdir mongrel2
> cd mongrel2
> fossil open ~/fossils/mongrel2.fossil
\end{Verbatim}
\end{code}

Make sure you do this in order (just like with every set of instructions you follow)
or else you'll get errors.  For example, if you don't make the \file{~/fossils} directory
then you'll get an error saying fossil can't open your file to clone.  Well, that's because
you \emph{didn't make the ~/fossils directory}.  Just pay attention and don't blame fossil
if you can't.


\subsection{Using the .zip File}

If you don't want to install fossil then you will need to get the .zip file
from our site and install it that way.  Since the code is new, you have \emph{no guarantees}
that this will work at all or that you will get a stable server.  We have not made an
official release, so using the .zip file is effectively a lame way of avoiding using
fossil.

\begin{enumerate}
\item Login as an anonymous user \href{http://mongrel2.org/login}{at mongrel2.org}.
\item Go to the \href{http://mongrel2.org/info/tip}{latest source tip}.
\item Click on the "ZIP archive" link to download the latest zip file of the source.
\item Unzip this file and cd into the resulting directory.
\end{enumerate}


\section{Building And Installing}

Once you have the source ready to go you can build it and then install it with
one command:  \shell{make all install}

There is no \shell{./configure} for Mongrel2 since we avoid too many OS specific
differences or shield those away with good feature checks in the code.

The end result of this should be:

\begin{enumerate}
\item Mongrel2 builds and compiles without errors.
\item All the unit tests run.\footnote{Please tell us about failures.}
\item The Python libraries that m2sh needs get installed.
\item The m2sh itself gets installed.
\item Finally, the mongrel2 binary gets installed.
\end{enumerate}

If any of these stages fail, then you can simply try to fix them and then
run:  \shell{make clean all install} which will do everything all over again.

The most common error is if you didn't install pip or distribute and insist
on using your own system's package management.  Do \emph{not} report bugs about
your own flavor of Linux's stupidity about versions and how it installs Python.
If you don't want to use the package system every other Python programmer uses,
then you're on your own.

\begin{aside}{Use Virtualenv If You Fear Pip}
For those who still want to use the package management but want to use pip and
distribute, I suggest you check out \href{http://pypi.python.org/pypi/virtualenv}{virtualenv}.
What virtualenv does is create a little "chroot install" of just the packages
you need using pip or distribute.  When you don't want them anymore, you can just
delete one directory and you're fine.
\end{aside}


\section{Testing The Installation}

When you are done, you probably want to make sure that it installed correctly.
There's a test configuration file in \file{tests/config.sqlite} that you can
use to try it out:

\begin{code}{First Test Run}
\begin{Verbatim}
> mkdir run
> mkdir logs
> mkdir tmp
> m2sh dump -db tests/config.sqlite
> m2sh start -db tests/config.sqlite -host localhost
\end{Verbatim}
\end{code}

That's it.  Just hit CTRL-c for now and we'll get into playing with this
setup later.

\section{Upgrading from trunk}
\begin{code}{Update your checkout}
\begin{Verbatim}
> cd mongrel2
> fossil pull http://mongrel2.org
> fossil update
\end{Verbatim}
\end{code}

Don't forget \shell{make clean all install}.


\section{Up Next}

You now should have a working Mongrel2 system installed and the m2sh configuration
interface ready to go.  In the rest of this manual we'll be simply learning how
to do more with Mongrel2, like making our own configs, writing handlers, and other
fun stuff.

